---
title: "Multiple trajectory graphical iterator"
author: "Phileas Dazeley Gaist"
date: "1/10/2022"
theme: cerulean
output:
  md_document:
    variant: gfm 
  pdf_document:
    highlight: default
    keep_tex: no
    fig_caption: yes
    latex_engine: pdflatex
fontsize: 11pt
geometry: margin=1in
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Graphical iterator program

I've written a simple graphical iterator program to help make graphical iterations
on the fly. Here it is:

I can't guarantee it'll work for all functions you throw at it, but then again, I 
can't think of any reason why it wouldn't. Let me know if you come across any or 
issues!

## Setup

```{r}
# set your initial condition and desired number of iterations:
x_0s <- c(3.43, 4.43, 7)
N <- 100

# set the iteration plot x axis range (lower and upper bounds):
x_min <- 0; x_max <- 8
y_min <- -2; y_max <- 8

use_custom_range_x <- FALSE
use_custom_range_y <- FALSE

# declare your function here:
func <- function(x){
  return(-2 * sin(x) + x) # function goes here
}
```

## The nitty-gritty

```{r}
get_function_data <- function(range = c(-1, 1), steps = 100){
  
  steps_multiplier <- (range[2]-range[1])/10 
  if(steps_multiplier < 1){steps_multiplier <- 1}
  # adds steps to get data for depending on the number of 10s 
  # in the specified plot x range
  
  x <- seq(from = range[1], to = range[2], length.out = steps * steps_multiplier)
  
  y <- array(dim = steps * steps_multiplier) 
  for(i in 1:length(x)){
    y[i] <- func(x[i])
    } 
  
  return(data.frame(x = x, y = y))
}

graphical_iterator <- function(x_0s, N = 100){ 
  
  segments <- data.frame()
  for(i in x_0s){
    
    start <- i 
    vert <- FALSE 
    
    x_0 <- rep(i,times=1+(N*2))
    xstarts <- c(start)
    ystarts <- c(y_min)
    xends <- c(start)
    yends <- c(func(start))
    
    # iteratively get the coordinates of the next segment points
    for(i in 1:(2 * N)) 
      # range = 2 * N because every step will be described by two segments
    {
      # if the last segment was vertical, the next must be horizontal
      if(vert){
        xstarts <- c(xstarts, start)
        ystarts <- c(ystarts, start)
        xends <- c(xends, start)
        yends <- c(yends, func(start)) 
        vert <- FALSE
      }
      else{
        xstarts <- c(xstarts, start)
        ystarts <- c(ystarts, func(start)) 
        xends <- c(xends, func(start)) 
        yends <- c(yends, func(start)) 
        vert <- TRUE
        start <- func(start) # update start value
      }
    }
    segments <- rbind(segments, data.frame(x_0s = x_0, xstarts, ystarts, xends, yends))
  }
  return(segments)
}

cobweb_trajects <- graphical_iterator(x_0s = x_0s, N = N)

if(use_custom_range_x == FALSE){
  x_min <- min(cobweb_trajects$xstarts); x_max <- max(cobweb_trajects$xends)
}
if(use_custom_range_y == FALSE){
  y_min <- min(cobweb_trajects$xstarts); y_max <- max(cobweb_trajects$xends)
}

plot_data <- get_function_data(range = c(x_min,x_max)) # gets the plotting data

get_function_iteration_trajectories <- function(x_0s, N = 100){
  
  trajectories <- data.frame()
  
  for(i in x_0s){
    x_t <- i
    
    x_0 <- rep(i,times=N+1)
    n <- 0:N
    
    trajectory <- c(x_t)
    
    for(t in 0:(N-1)){
      x_t <- func(x_t)
      trajectory <- c(trajectory, x_t) # add x_t_1's value to the trajectory vector
    }
    trajectories <- rbind(trajectories, data.frame(x_0s = x_0, ns = n, trajectories = trajectory))
  }
  return(trajectories)
}

trajectories <- get_function_iteration_trajectories(x_0s = x_0s, N = N)
```

## Plots

**Graphical iteration plot:**

```{r}
plot_data %>% 
  ggplot(aes(x, y)) +
  geom_line(colour = "black") +
  geom_abline(linetype = "dashed") + 
  geom_segment(data = cobweb_trajects, aes(x = xstarts, y = ystarts, xend = xends, 
                                          yend = yends, colour=as.factor(x_0s))) +
  coord_cartesian(xlim = c(x_min, x_max), ylim = c(y_min, y_max)) 
```

TODO IN THE FUTURE: 

- Colour segments based on distance to fixed points. 
- Colour segments based on distance to other segments.

**Iteration trajectory time series plot**

```{r}
trajectories

# trajectory plot
trajectories %>% 
  ggplot(aes(ns, trajectories, colour = as.factor(x_0s))) +
  geom_line() + labs(x="n")
```
```{r}
# TODO
# average distances between trajectories plot

# mean(dist(c(1:10))) # mean of distances between whole numbers from 1 to 10 (example)

# trajectories$x_0s <- paste0('x_0_', trajectories$x_0s)
# trajectories_wide <- trajectories %>% 
#   pivot_wider(names_from = x_0s, values_from = trajectories) %>% 
#   mutate(distance = (x_0_2.01-x_0_2))
# 
# trajectories_wide
# 
# trajectories_wide %>% 
#   ggplot(aes(ns, distance)) +
#   geom_point() + geom_line() +
#   labs(x="n", y="distance between trajectories")
```

